---
title: "Activity 02"
author: "CG2CG2"
date: "`r Sys.Date()`"
output: 
  pdf_document:
          latex_engine: xelatex
---

## Data Generation

```{r setup, message=FALSE, warning=FALSE, echo=TRUE}
# Set seed for reproducibility
set.seed(22)

# Generate dataset (60 observations)
activity_data <- data.frame(
  StudyHours  = rnorm(60, mean = 6,  sd = 2),
  Attendance  = rnorm(60, mean = 85, sd = 8),
  Score       = rnorm(60, mean = 78, sd = 7),
  Gender      = sample(c("Male", "Female"), 60, replace = TRUE),
  Group       = sample(c("Group 1", "Group 2", "Group 3"), 60, replace = TRUE)
)

activity_data$Gender <- factor(activity_data$Gender)
activity_data$Group  <- factor(activity_data$Group)
```

---

# A. Summary Statistics

## A1. Compute Descriptive Statistics

```{r}
summary(activity_data)
```

## A2. Measure Variability

```{r}
sapply(activity_data[c("StudyHours","Attendance","Score")], sd)
```

## A3. Check Minimum and Maximum Values

```{r}
sapply(activity_data[c("StudyHours","Attendance","Score")], range)
```

### Interpretation

- The average values of the numerical variables are:
  - **StudyHours:** 6.249  
  - **Attendance:** 84.22  
  - **Score:** 76.69  

- The variable with the greatest variability is:
  - **Attendance**, as it has the highest standard deviation (7.87), indicating the widest spread of values.

- Any unusual minimum or maximum values observed:
  - Attendance has a maximum value of **101.97**, which slightly exceeds 100. This is possible because the data were generated using a normal distribution without upper bounds. No extreme outliers are observed in the other variables.


---

# B. Data Structure

```{r}
str(activity_data)
```

### Answers

- The object class of `activity_data` is **data.frame**.
- The dataset contains **60 observations and 5 variables**.
- The variables are: StudyHours, Attendance, Score, Gender, and Group.
- StudyHours, Attendance, and Score are **numeric variables**.
- Gender and Group are **categorical variables (factors)**.

---

# C. Probability Estimation

## Estimate Probability that Score > 85

```{r}
mean(activity_data$Score > 85)
```

### Interpretation

It is empirical because it was calculated using the observed sample data rather than a theoretical probability formula based on the population distribution.

---

# D. Distribution of Scores

```{r}
hist(activity_data$Score,
     main = "Histogram of Score",
     xlab = "Score",
     col = "lightblue",
     border = "black")
```

### Interpretation

- The histogram of Score appears approximately bell-shaped and reasonably symmetric, indicating that the distribution is roughly normal.
- Yes, assuming a normal distribution is reasonable because the data were generated using a normal distribution (`rnorm`), and the observed shape aligns with that expectation.

---

# E. Expectation and Variance of Score

Theoretical expectation (mean) = 78
Theoretical variance = 49

## Sample Estimates

```{r}
mean(activity_data$Score)
var(activity_data$Score)
```

### Explanation

The sample mean (76.69) and sample variance (39.66) differ from theoretical values because they are **estimates based on a finite sample**, not exact population parameters.

---

# F. Correlation Analysis

```{r}
cor(activity_data[c("StudyHours","Attendance","Score")])
```

### Interpretation

- StudyHours and Attendance: **r = 0.061** → negligible positive relationship.
- StudyHours and Score: **r = -0.065** → negligible negative relationship.
- Attendance and Score: **r = 0.227** → low positive relationship.

Overall, there are **no strong linear relationships** among the numerical variables.  
Correlation does not imply causation.

---

# G. One-Way ANOVA (Score by Group)

```{r}
oneway.test(Score ~ Group, data = activity_data)
```

### Interpretation

* p-value: 0.2379
* Decision at α = 0.05: since p > 0.05, we **do not reject the null hypothesis**.
* Conclusion: **no statistically significant difference** in mean Score among the three Groups.

---

# H. Chi-Square Test (Gender vs Group)

```{r}
tab <- table(activity_data$Gender, activity_data$Group)
tab
chisq.test(tab)
```

### Interpretation

* p-value: 0.5258
* Conclusion regarding association:
- There is **no significant association** between Gender and Group.
- Gender and Group appear to be independent in this dataset.




