---
title: "CG2CG2-ACTIVITY01"
author: "CG2CG2"
output:
  pdf_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### PART 1: LOAD DATA
```{r}
setwd("C:/Users/sethd/Documents/CS3203N/CG2CG2_ACT1")

library(dplyr)

edata <- read.csv("employee_data.csv")

edata
head(edata)

colSums(is.na(edata))
edata %>% filter(is.na(Performance))
```
Based on the first six rows of the dataset, the *Performance* variable contains two missing values (NA).

### PART 2: SELECT AND FILTER

#### Part 2(a): Select Name, Department, Salary
```{r}
edata_sel <- edata %>% select(Name, Department, Salary)
edata_sel

edata_sel_base <- data.frame(
  Name = edata$Name,
  Department = edata$Department,
  Salary = edata$Salary
)
edata_sel_base
```
#### Part 2(b): Filter Salary > 55,000
```{r}
edata_high_salary <- edata %>% filter(Salary > 55000)
edata_high_salary
```
### PART 3: MUTATE AND ARRANGE
#### Part 3(a): Create Seniority column
```{r}
edata_senior <- edata %>% mutate(Seniority = YearsWorked > 5)
edata_senior
```
#### Part 3(b): Arrange by Performance (descending)
```{r}
edata_sorted_perf <- edata %>% arrange(desc(Performance))
edata_sorted_perf
```
#### Part 3(c): Describe the last output
Employees are sorted from highest to lowest performance rating.
Rows with missing performance values appear at the bottom.

### PART 4: GROUPED OPERATIONS
#### Part 4(a): Average salary by Department
```{r}
dept_avg_salary <- edata %>%
  group_by(Department) %>%
  summarise(MeanSalary = mean(Salary), .groups = "drop")
dept_avg_salary

aggregate(Salary ~ Department, data = edata, mean)
```
#### Part 4(b): Count employees per Department
```{r}
dept_counts <- edata %>%
  group_by(Department) %>%
  summarise(EmployeeCount = n(), .groups = "drop")
dept_counts
```
#### Part 4(c): Interpretation
The IT department has the highest average salary.
The IT and Sales departments have the most employees, with three employees each.







